---
title: "Investigating Suicide Rate"
author: "490473041 ,490474923, 490387829, 490411489"
date: "March 22, 2019"
output: ioslides_presentation


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# LOAD DATA v1 - uncomment the link below to: load data direct from html
setwd("C:/Users/anosh/Documents/Anosh's Data/!_Uni_!/DATA1001 - Intro to Data/Assignment 1")
su_rate1 = read.csv("Data/who_suicide_statistics_NOBLANKS.csv")
su_rate2 = read.csv("Data/Tidy Data Set - Suicide Rates (REVISED).csv")

# LOAD DATA v2 - uncomment the link below to: load data from local file
#cars = read.csv("dataset file location")
library(plotly)
```

## Executive Summary

- Suicide is a growing problem in many places around the globe in the 21st century. 
- Delineating a correlation between socio-economic and health factors and suicide rate through the comparison of gross domestic product, age, sex, and unemployment rate
- Hence exploring the trends in the data to understand why the trends found exist and what is potentially driving the changes


## IDA

- Suicide is a complex issue that contributes to over 800,000 deaths annually
- Comparing socio economic and health factors such as GDP at PPP, age, sex, and unemployment rate to suicide rate allows us to understand whether these factors are correlated with suicide rate
- The data was sourced from the WHO and the World Bank, two reputable sources, to ensure the validity of the data


## Investigating the Relationship Between Income ($ PPP) and Suicide Rate

- According to the WHO (2018), low and middle income represent 79% of completed suicide.  
- Increased levels of income past a certain point may result in lower life satisfaction (Jebb, Tay, Diener, & Oishi, 2018) 
- Loss of pleasure in life is a psychosocial state which increases risk of completing suicide (Cucher, & Chehill, 2012). 
- Measurement for Income: GDP per capita at PPP

## 

```{r, message = FALSE}


GDP2000 <- filter(su_rate2, su_rate2$Year == "2000" & su_rate2$GDP.Per.Capita....PPP. > 0)

x <- list(
  title = "Income ($ PPP)"
)

y <- list(
  title = "Suicide Rate (per 100,000)"
)

p1 <- plot_ly(GDP2000, x = ~GDP.Per.Capita....PPP., y = ~Suicide.Rate, color = ~Region, type = "scatter",
             text = ~paste("Region: ",Region, '<br>Country:', Country,'<br>Suicide Rate:', Suicide.Rate, '<br>Income ($ PPP):', GDP.Per.Capita....PPP., '<br>Gender:', Gender)) %>%

    layout(
    title = "Suicide Rate vs. Income($ PPP) Across Different Countries and Regions (2000)"
    , xaxis = x, yaxis = y) 
p1


```

##

```{r, message = FALSE}
library(plotly)

GDP2005 <- filter(su_rate2, su_rate2$Year == "2005" & su_rate2$GDP.Per.Capita....PPP. > 0)

x <- list(
  title = "Income ($ PPP)"
)

y <- list(
  title = "Suicide Rate (per 100,000)"
)

p1 <- plot_ly(GDP2005, x = ~GDP.Per.Capita....PPP., y = ~Suicide.Rate, color = ~Region, type = "scatter",
             text = ~paste("Region: ",Region, '<br>Country:', Country, '<br>Suicide Rate:', Suicide.Rate, '<br>Income ($ PPP):', GDP.Per.Capita....PPP., '<br>Gender:', Gender)) %>%

    layout(
    title = "Suicide Rate vs. Income($ PPP) Across Different Countries and Regions (2005)"
    , xaxis = x, yaxis = y) 
p1


```

##

```{r, message = FALSE}
library(plotly)

GDP2010 <- filter(su_rate2, su_rate2$Year == "2010" & su_rate2$GDP.Per.Capita....PPP. > 0)

x <- list(
  title = "Income ($ PPP)"
)

y <- list(
  title = "Suicide Rate (per 100,000)"
)

p1 <- plot_ly(GDP2010, x = ~GDP.Per.Capita....PPP., y = ~Suicide.Rate, color = ~Region, type = "scatter",
             text = ~paste("Region: ",Region, '<br>Country:', Country, '<br>Suicide Rate:', Suicide.Rate, '<br>Income ($ PPP):', GDP.Per.Capita....PPP., '<br>Gender:', Gender)) %>%

    layout(
    title = "Suicide Rate vs. Income($ PPP) Across Different Countries and Regions (2010)"
    , xaxis = x, yaxis = y) 
p1


```

##

```{r, message = FALSE}
library(plotly)

GDP2015 <- filter(su_rate2, su_rate2$Year == "2015" & su_rate2$GDP.Per.Capita....PPP. > 0)

x <- list(
  title = "Income ($ PPP)"
)

y <- list(
  title = "Suicide Rate (per 100,000)"
)

p1 <- plot_ly(GDP2015, x = ~GDP.Per.Capita....PPP., y = ~Suicide.Rate, color = ~Region, type = "scatter",
             text = ~paste("Region: ",Region, '<br>Country:', Country, '<br>Suicide Rate:', Suicide.Rate, '<br>Income ($ PPP):', GDP.Per.Capita....PPP., '<br>Gender:', Gender)) %>%

    layout(
    title = "Suicide Rate vs. Income($ PPP) Across Different Countries and Regions (2015)"
    , xaxis = x, yaxis = y) 
p1


```

## Investigating the Relationship Between Unemployment and Suicide Rate

- A causal link between suicide and unemployment as the impact of being unemployed 
- Impact of being unemployed is not simply the economic pressure of a reduced income but also, the loss of status and structure which work provides
- Social stigma attached to the loss of status adds greater pressure on unemployed individuals

## Investigating the Relationship Between Unemployment and Suicide Rate (Cont.)
- Suicide rates peaked in 1930's during Great Depression (Social Science & Medicine, Volume 36 Issue 6 Pages 715-841 (March 1993))
- 45,000 (1 in 5) of suicides attributed to unemployment (Carlos Nordt and colleagues at the University of Zurich)
- 20-30% rise in unemployment led to 12% rise in suicide between 2000-15 (Our dataset and Dr Carlos Nordt in Lancet Psychiatry Volume 2, Issue 3, P239-245, March 01, 2015)


##

```{r, echo = FALSE, message=FALSE}


UR2000 <- filter(su_rate2, su_rate2$Year == "2000" & su_rate2$Unemployment.Rate > 0)

## 
x <- list(
  title = "Unemployment Rate (%)"
)

y <- list(
  title = "Suicide Rate (per 100,000)"
)

p1 <- plot_ly(UR2000, x = ~Unemployment.Rate, y = ~Suicide.Rate, color = ~Region, type = "scatter",
             text = ~paste("Region: ",Region, '<br>Country:', Country, '<br>Suicide Rate:', Suicide.Rate, '<br>Unemployment Rate (%):', Unemployment.Rate, '<br>Gender:', Gender)) %>%

    layout(
    title = "Suicide Rate vs. Unemployment Rate (%) Across Different Countries and Regions (2000)"
    , xaxis = x, yaxis = y) 
p1


```

##

```{r, message=FALSE}
library(plotly)

UR2005 <- filter(su_rate2, su_rate2$Year == "2005" & su_rate2$Unemployment.Rate > 0)

x <- list(
  title = "Unemployment Rate (%)"
)

y <- list(
  title = "Suicide Rate (per 100,000)"
)

p1 <- plot_ly(UR2005, x = ~Unemployment.Rate, y = ~Suicide.Rate, color = ~Region, type = "scatter",
             text = ~paste("Region: ",Region, '<br>Country:', Country, '<br>Suicide Rate:', Suicide.Rate, '<br>Unemployment Rate (%):', Unemployment.Rate, '<br>Gender:', Gender)) %>%

    layout(
    title = "Suicide Rate vs. Unemployment Rate (%) Across Different Countries and Regions (2005)"
    , xaxis = x, yaxis = y) 
p1


```

##


```{r, message=FALSE}
library(plotly)

UR2010 <- filter(su_rate2, su_rate2$Year == "2010" & su_rate2$Unemployment.Rate > 0)

x <- list(
  title = "Unemployment Rate (%)"
)

y <- list(
  title = "Suicide Rate (per 100,000)"
)

p1 <- plot_ly(UR2010, x = ~Unemployment.Rate, y = ~Suicide.Rate, color = ~Region, type = "scatter",
             text = ~paste("Region: ",Region, '<br>Country:', Country, '<br>Suicide Rate:', Suicide.Rate, '<br>Unemployment Rate (%):', Unemployment.Rate, '<br>Gender:', Gender)) %>%

    layout(
    title = "Suicide Rate vs. Unemployment Rate (%) Across Different Countries and Regions (2010)"
    , xaxis = x, yaxis = y) 
p1


```

##

```{r, message=FALSE}
library(plotly)

UR2015 <- filter(su_rate2, su_rate2$Year == "2015" & su_rate2$Unemployment.Rate > 0)

x <- list(
  title = "Unemployment Rate (%)"
)

y <- list(
  title = "Suicide Rate (per 100,000)"
)

p1 <- plot_ly(UR2015, x = ~Unemployment.Rate, y = ~Suicide.Rate, color = ~Region, type = "scatter",
             text = ~paste("Region: ",Region, '<br>Country:', Country, '<br>Suicide Rate:', Suicide.Rate, '<br>Unemployment Rate (%):', Unemployment.Rate, '<br>Gender:', Gender)) %>%

    layout(
    title = "Suicide Rate vs. Unemployment Rate (%) Across Different Countries and Regions (2015)"
    , xaxis = x, yaxis = y) 
p1


```

## Investigating the Relationship Between Age, Sex and Suicide Rate

- The US is a distinct case as it is the only country with a consistently increasing suicide rate
- The average US male suicide is 23.36 compared to the world male average of 16.63, suggesting an important issue that America must address
- This high rate is attributed not to mental health issues but to relationship, substance, physical health, and monetary issues
- The world average suicide rate for women is 4.5 while the corresponding rate for men is 16.6. This is 3.7 times high than the women's rate
- This trend of the men's rate being higher than the women's translated across the vast majority of countries in the data

##

```{r}

library(plotly)

a <-filter(su_rate1, su_rate1$year == "2000" &  su_rate1$sex == "male" & su_rate1$age =="5-14 years") 
a1 <- mean(a$rate)
a2 <-filter(su_rate1, su_rate1$year == "2000" & su_rate1$sex == "female" & su_rate1$age =="5-14 years")
a3 <- mean(a2$rate)

b <-filter(su_rate1, su_rate1$year == "2000" & su_rate1$sex == "male" & su_rate1$age =="15-24 years") 
b1 <- mean(b$rate)
b2 <-filter(su_rate1, su_rate1$year == "2000" & su_rate1$sex == "female" & su_rate1$age =="15-24 years")
b3 <- mean(b2$rate)

c <-filter(su_rate1, su_rate1$year == "2000" & su_rate1$sex == "male" & su_rate1$age =="25-34 years") 
c1 <- mean(c$rate)
c2 <-filter(su_rate1, su_rate1$year == "2000" & su_rate1$sex == "female" & su_rate1$age =="25-34 years")
c3 <- mean(c2$rate)

d <-filter(su_rate1, su_rate1$year == "2000" & su_rate1$sex == "male" & su_rate1$age =="35-54 years") 
d1 <- mean(d$rate)
d2 <-filter(su_rate1, su_rate1$year == "2000" & su_rate1$sex == "female" & su_rate1$age =="35-54 years")
d3 <- mean(d2$rate)

e <-filter(su_rate1, su_rate1$year == "2000" & su_rate1$sex == "male" & su_rate1$age =="55-74 years") 
e1 <- mean(e$rate)
e2 <-filter(su_rate1, su_rate1$year == "2000" & su_rate1$sex == "female" & su_rate1$age =="55-74 years")
e3 <- mean(e2$rate)

f <-filter(su_rate1, su_rate1$year == "2000" & su_rate1$sex == "male" & su_rate1$age =="75+ years") 
f1 <- mean(f$rate)
f2 <-filter(su_rate1, su_rate1$year == "2000" & su_rate1$sex == "female" & su_rate1$age =="75+ years")
f3 <- mean(f2$rate)

table <- data.frame(x = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"),
                    y = c(a1, b1, c1, d1, e1, f1))
table$x <- factor(table$x, levels = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"))


p1 <- plot_ly(
  data = table,
  x = ~x,
  y = ~y, 
  type = "bar",
  name = "Male",
  color = I("dodgerblue4")
 
  
) %>%
  
  add_trace(
    y = c(a3, b3, c3, d3, e3, f3),
    name = "Female",
    color = I("firebrick4")
      ) %>%
  
  layout(
    title = "World Average Suicide Rate Across Different Age Categories (2000)", 
    yaxis = list(
      title = "Suicide Rate (per 100,000)"),
    xaxis = list(
      title = "Age"),
    
      barmode = "group"
    )
  

p1

```

##

```{r}

library(plotly)

a <-filter(su_rate1, su_rate1$year == "2005" & su_rate1$sex == "male" & su_rate1$age =="5-14 years") 
a1 <- mean(a$rate)
a2 <-filter(su_rate1, su_rate1$year == "2005" & su_rate1$sex == "female" & su_rate1$age =="5-14 years")
a3 <- mean(a2$rate)

b <-filter(su_rate1, su_rate1$year == "2005" & su_rate1$sex == "male" & su_rate1$age =="15-24 years") 
b1 <- mean(b$rate)
b2 <-filter(su_rate1, su_rate1$year == "2005" & su_rate1$sex == "female" & su_rate1$age =="15-24 years")
b3 <- mean(b2$rate)

c <-filter(su_rate1, su_rate1$year == "2005" & su_rate1$sex == "male" & su_rate1$age =="25-34 years") 
c1 <- mean(c$rate)
c2 <-filter(su_rate1, su_rate1$year == "2005" & su_rate1$sex == "female" & su_rate1$age =="25-34 years")
c3 <- mean(c2$rate)

d <-filter(su_rate1, su_rate1$year == "2005" & su_rate1$sex == "male" & su_rate1$age =="35-54 years") 
d1 <- mean(d$rate)
d2 <-filter(su_rate1, su_rate1$year == "2005" & su_rate1$sex == "female" & su_rate1$age =="35-54 years")
d3 <- mean(d2$rate)

e <-filter(su_rate1, su_rate1$year == "2005" & su_rate1$sex == "male" & su_rate1$age =="55-74 years") 
e1 <- mean(e$rate)
e2 <-filter(su_rate1, su_rate1$year == "2005" & su_rate1$sex == "female" & su_rate1$age =="55-74 years")
e3 <- mean(e2$rate)

f <-filter(su_rate1, su_rate1$year == "2005" & su_rate1$sex == "male" & su_rate1$age =="75+ years") 
f1 <- mean(f$rate)
f2 <-filter(su_rate1, su_rate1$year == "2005" & su_rate1$sex == "female" & su_rate1$age =="75+ years")
f3 <- mean(f2$rate)

table <- data.frame(x = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"),
                    y = c(a1, b1, c1, d1, e1, f1))
table$x <- factor(table$x, levels = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"))


p1 <- plot_ly(
  data = table,
  x = ~x,
  y = ~y, 
  type = "bar",
  name = "Male",
  color = I("dodgerblue4")
 
  
) %>%
  
  add_trace(
    y = c(a3, b3, c3, d3, e3, f3),
    name = "Female",
    color = I("firebrick4")
      ) %>%
  
  layout(
    title = "World Average Suicide Rate Across Different Age Categories (2005)", 
    yaxis = list(
      title = "Suicide Rate (per 100,000)"),
    xaxis = list(
      title = "Age"),
    
      barmode = "group"
    )
  

p1

```

##

```{r}

library(plotly)

a <-filter(su_rate1, su_rate1$year == "2010" & su_rate1$sex == "male" & su_rate1$age =="5-14 years") 
a1 <- mean(a$rate)
a2 <-filter(su_rate1, su_rate1$year == "2010" & su_rate1$sex == "female" & su_rate1$age =="5-14 years")
a3 <- mean(a2$rate)

b <-filter(su_rate1, su_rate1$year == "2010" & su_rate1$sex == "male" & su_rate1$age =="15-24 years") 
b1 <- mean(b$rate)
b2 <-filter(su_rate1, su_rate1$year == "2010" & su_rate1$sex == "female" & su_rate1$age =="15-24 years")
b3 <- mean(b2$rate)

c <-filter(su_rate1, su_rate1$year == "2010" & su_rate1$sex == "male" & su_rate1$age =="25-34 years") 
c1 <- mean(c$rate)
c2 <-filter(su_rate1, su_rate1$year == "2010" & su_rate1$sex == "female" & su_rate1$age =="25-34 years")
c3 <- mean(c2$rate)

d <-filter(su_rate1, su_rate1$year == "2010" & su_rate1$sex == "male" & su_rate1$age =="35-54 years") 
d1 <- mean(d$rate)
d2 <-filter(su_rate1, su_rate1$year == "2010" & su_rate1$sex == "female" & su_rate1$age =="35-54 years")
d3 <- mean(d2$rate)

e <-filter(su_rate1, su_rate1$year == "2010" & su_rate1$sex == "male" & su_rate1$age =="55-74 years") 
e1 <- mean(e$rate)
e2 <-filter(su_rate1, su_rate1$year == "2010" & su_rate1$sex == "female" & su_rate1$age =="55-74 years")
e3 <- mean(e2$rate)

f <-filter(su_rate1, su_rate1$year == "2010" & su_rate1$sex == "male" & su_rate1$age =="75+ years") 
f1 <- mean(f$rate)
f2 <-filter(su_rate1, su_rate1$year == "2010" & su_rate1$sex == "female" & su_rate1$age =="75+ years")
f3 <- mean(f2$rate)

table <- data.frame(x = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"),
                    y = c(a1, b1, c1, d1, e1, f1))
table$x <- factor(table$x, levels = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"))


p1 <- plot_ly(
  data = table,
  x = ~x,
  y = ~y, 
  type = "bar",
  name = "Male",
  color = I("dodgerblue4")
 
  
) %>%
  
  add_trace(
    y = c(a3, b3, c3, d3, e3, f3),
    name = "Female",
    color = I("firebrick4")
      ) %>%
  
  layout(
    title = "World Average Suicide Rate Across Different Age Categories (2010)", 
    yaxis = list(
      title = "Suicide Rate (per 100,000)"),
    xaxis = list(
      title = "Age"),
    
      barmode = "group"
    )
  

p1

```

##

```{r}

library(plotly)

a <-filter(su_rate1, su_rate1$year == "2015" & su_rate1$sex == "male" & su_rate1$age =="5-14 years") 
a1 <- mean(a$rate)
a2 <-filter(su_rate1, su_rate1$year == "2015" & su_rate1$sex == "female" & su_rate1$age =="5-14 years")
a3 <- mean(a2$rate)

b <-filter(su_rate1, su_rate1$year == "2015" & su_rate1$sex == "male" & su_rate1$age =="15-24 years") 
b1 <- mean(b$rate)
b2 <-filter(su_rate1, su_rate1$year == "2015" & su_rate1$sex == "female" & su_rate1$age =="15-24 years")
b3 <- mean(b2$rate)

c <-filter(su_rate1, su_rate1$year == "2015" & su_rate1$sex == "male" & su_rate1$age =="25-34 years") 
c1 <- mean(c$rate)
c2 <-filter(su_rate1, su_rate1$year == "2015" & su_rate1$sex == "female" & su_rate1$age =="25-34 years")
c3 <- mean(c2$rate)

d <-filter(su_rate1, su_rate1$year == "2015" & su_rate1$sex == "male" & su_rate1$age =="35-54 years") 
d1 <- mean(d$rate)
d2 <-filter(su_rate1, su_rate1$year == "2015" & su_rate1$sex == "female" & su_rate1$age =="35-54 years")
d3 <- mean(d2$rate)

e <-filter(su_rate1, su_rate1$year == "2015" & su_rate1$sex == "male" & su_rate1$age =="55-74 years") 
e1 <- mean(e$rate)
e2 <-filter(su_rate1, su_rate1$year == "2015" & su_rate1$sex == "female" & su_rate1$age =="55-74 years")
e3 <- mean(e2$rate)

f <-filter(su_rate1, su_rate1$year == "2015" & su_rate1$sex == "male" & su_rate1$age =="75+ years") 
f1 <- mean(f$rate)
f2 <-filter(su_rate1, su_rate1$year == "2015" & su_rate1$sex == "female" & su_rate1$age =="75+ years")
f3 <- mean(f2$rate)

table <- data.frame(x = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"),
                    y = c(a1, b1, c1, d1, e1, f1))
table$x <- factor(table$x, levels = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"))


p1 <- plot_ly(
  data = table,
  x = ~x,
  y = ~y, 
  type = "bar",
  name = "Male",
  color = I("dodgerblue4")
 
  
) %>%
  
  add_trace(
    y = c(a3, b3, c3, d3, e3, f3),
    name = "Female",
    color = I("firebrick4")
      ) %>%
  
  layout(
    title = "World Average Suicide Rate Across Different Age Categories (2015)", 
    yaxis = list(
      title = "Suicide Rate (per 100,000)"),
    xaxis = list(
      title = "Age"),
    
      barmode = "group"
    )
  

p1

```

##

```{r}
library(plotly)

a <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2000" & su_rate1$sex == "male" & su_rate1$age =="5-14 years") 
a1 <- mean(a$rate)
a2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2000" & su_rate1$sex == "female" & su_rate1$age =="5-14 years")
a3 <- mean(a2$rate)

b <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2000" & su_rate1$sex == "male" & su_rate1$age =="15-24 years") 
b1 <- mean(b$rate)
b2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2000" & su_rate1$sex == "female" & su_rate1$age =="15-24 years")
b3 <- mean(b2$rate)

c <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2000" & su_rate1$sex == "male" & su_rate1$age =="25-34 years") 
c1 <- mean(c$rate)
c2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2000" & su_rate1$sex == "female" & su_rate1$age =="25-34 years")
c3 <- mean(c2$rate)

d <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2000" & su_rate1$sex == "male" & su_rate1$age =="35-54 years") 
d1 <- mean(d$rate)
d2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2000" & su_rate1$sex == "female" & su_rate1$age =="35-54 years")
d3 <- mean(d2$rate)

e <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2000" & su_rate1$sex == "male" & su_rate1$age =="55-74 years") 
e1 <- mean(e$rate)
e2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2000" & su_rate1$sex == "female" & su_rate1$age =="55-74 years")
e3 <- mean(e2$rate)

f <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2000" & su_rate1$sex == "male" & su_rate1$age =="75+ years") 
f1 <- mean(f$rate)
f2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2000" & su_rate1$sex == "female" & su_rate1$age =="75+ years")
f3 <- mean(f2$rate)

table <- data.frame(x = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"),
                    y = c(a1, b1, c1, d1, e1, f1))
table$x <- factor(table$x, levels = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"))


p1 <- plot_ly(
  data = table,
  x = ~x,
  y = ~y, 
  type = "bar",
  name = "Male",
  color = I("dodgerblue4")
 
  
) %>%
  
  add_trace(
    y = c(a3, b3, c3, d3, e3, f3),
    name = "Female",
    color = I("firebrick4")
      ) %>%
  
  layout(
    title = "US Average Suicide Rate Across Different Age Categories (2000)", 
    yaxis = list(
      title = "Suicide Rate (per 100,000)"),
    xaxis = list(
      title = "Age"),
    
      barmode = "group"
    )
  

p1

```

##

```{r}
library(plotly)

a <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2005" & su_rate1$sex == "male" & su_rate1$age =="5-14 years") 
a1 <- mean(a$rate)
a2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2005" & su_rate1$sex == "female" & su_rate1$age =="5-14 years")
a3 <- mean(a2$rate)

b <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2005" & su_rate1$sex == "male" & su_rate1$age =="15-24 years") 
b1 <- mean(b$rate)
b2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2005" & su_rate1$sex == "female" & su_rate1$age =="15-24 years")
b3 <- mean(b2$rate)

c <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2005" & su_rate1$sex == "male" & su_rate1$age =="25-34 years") 
c1 <- mean(c$rate)
c2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2005" & su_rate1$sex == "female" & su_rate1$age =="25-34 years")
c3 <- mean(c2$rate)

d <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2005" & su_rate1$sex == "male" & su_rate1$age =="35-54 years") 
d1 <- mean(d$rate)
d2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2005" & su_rate1$sex == "female" & su_rate1$age =="35-54 years")
d3 <- mean(d2$rate)

e <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2005" & su_rate1$sex == "male" & su_rate1$age =="55-74 years") 
e1 <- mean(e$rate)
e2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2005" & su_rate1$sex == "female" & su_rate1$age =="55-74 years")
e3 <- mean(e2$rate)

f <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2005" & su_rate1$sex == "male" & su_rate1$age =="75+ years") 
f1 <- mean(f$rate)
f2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2005" & su_rate1$sex == "female" & su_rate1$age =="75+ years")
f3 <- mean(f2$rate)

table <- data.frame(x = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"),
                    y = c(a1, b1, c1, d1, e1, f1))
table$x <- factor(table$x, levels = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"))


p1 <- plot_ly(
  data = table,
  x = ~x,
  y = ~y, 
  type = "bar",
  name = "Male",
  color = I("dodgerblue4")
 
  
) %>%
  
  add_trace(
    y = c(a3, b3, c3, d3, e3, f3),
    name = "Female",
    color = I("firebrick4")
      ) %>%
  
  layout(
    title = "US Average Suicide Rate Across Different Age Categories (2005)", 
    yaxis = list(
      title = "Suicide Rate (per 100,000)"),
    xaxis = list(
      title = "Age"),
    
      barmode = "group"
    )
  

p1

```

##

```{r}
library(plotly)

a <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2010" & su_rate1$sex == "male" & su_rate1$age =="5-14 years") 
a1 <- mean(a$rate)
a2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2010" & su_rate1$sex == "female" & su_rate1$age =="5-14 years")
a3 <- mean(a2$rate)

b <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2010" & su_rate1$sex == "male" & su_rate1$age =="15-24 years") 
b1 <- mean(b$rate)
b2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2010" & su_rate1$sex == "female" & su_rate1$age =="15-24 years")
b3 <- mean(b2$rate)

c <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2010" & su_rate1$sex == "male" & su_rate1$age =="25-34 years") 
c1 <- mean(c$rate)
c2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2010" & su_rate1$sex == "female" & su_rate1$age =="25-34 years")
c3 <- mean(c2$rate)

d <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2010" & su_rate1$sex == "male" & su_rate1$age =="35-54 years") 
d1 <- mean(d$rate)
d2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2010" & su_rate1$sex == "female" & su_rate1$age =="35-54 years")
d3 <- mean(d2$rate)

e <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2010" & su_rate1$sex == "male" & su_rate1$age =="55-74 years") 
e1 <- mean(e$rate)
e2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2010" & su_rate1$sex == "female" & su_rate1$age =="55-74 years")
e3 <- mean(e2$rate)

f <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2010" & su_rate1$sex == "male" & su_rate1$age =="75+ years") 
f1 <- mean(f$rate)
f2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2010" & su_rate1$sex == "female" & su_rate1$age =="75+ years")
f3 <- mean(f2$rate)

table <- data.frame(x = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"),
                    y = c(a1, b1, c1, d1, e1, f1))
table$x <- factor(table$x, levels = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"))


p1 <- plot_ly(
  data = table,
  x = ~x,
  y = ~y, 
  type = "bar",
  name = "Male",
  color = I("dodgerblue4")
 
  
) %>%
  
  add_trace(
    y = c(a3, b3, c3, d3, e3, f3),
    name = "Female",
    color = I("firebrick4")
      ) %>%
  
  layout(
    title = "US Average Suicide Rate Across Different Age Categories (2010)", 
    yaxis = list(
      title = "Suicide Rate (per 100,000)"),
    xaxis = list(
      title = "Age"),
    
      barmode = "group"
    )
  

p1

```

##

```{r}
library(plotly)

a <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2015" & su_rate1$sex == "male" & su_rate1$age =="5-14 years") 
a1 <- mean(a$rate)
a2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2015" & su_rate1$sex == "female" & su_rate1$age =="5-14 years")
a3 <- mean(a2$rate)

b <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2015" & su_rate1$sex == "male" & su_rate1$age =="15-24 years") 
b1 <- mean(b$rate)
b2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2015" & su_rate1$sex == "female" & su_rate1$age =="15-24 years")
b3 <- mean(b2$rate)

c <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2015" & su_rate1$sex == "male" & su_rate1$age =="25-34 years") 
c1 <- mean(c$rate)
c2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2015" & su_rate1$sex == "female" & su_rate1$age =="25-34 years")
c3 <- mean(c2$rate)

d <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2015" & su_rate1$sex == "male" & su_rate1$age =="35-54 years") 
d1 <- mean(d$rate)
d2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2015" & su_rate1$sex == "female" & su_rate1$age =="35-54 years")
d3 <- mean(d2$rate)

e <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2015" & su_rate1$sex == "male" & su_rate1$age =="55-74 years") 
e1 <- mean(e$rate)
e2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2015" & su_rate1$sex == "female" & su_rate1$age =="55-74 years")
e3 <- mean(e2$rate)

f <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2015" & su_rate1$sex == "male" & su_rate1$age =="75+ years") 
f1 <- mean(f$rate)
f2 <-filter(su_rate1, su_rate1$country == "United States of America" & su_rate1$year == "2015" & su_rate1$sex == "female" & su_rate1$age =="75+ years")
f3 <- mean(f2$rate)

table <- data.frame(x = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"),
                    y = c(a1, b1, c1, d1, e1, f1))
table$x <- factor(table$x, levels = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"))


p1 <- plot_ly(
  data = table,
  x = ~x,
  y = ~y, 
  type = "bar",
  name = "Male",
  color = I("dodgerblue4")
 
  
) %>%
  
  add_trace(
    y = c(a3, b3, c3, d3, e3, f3),
    name = "Female",
    color = I("firebrick4")
      ) %>%
  
  layout(
    title = "US Average Suicide Rate Across Different Age Categories (2015)", 
    yaxis = list(
      title = "Suicide Rate (per 100,000)"),
    xaxis = list(
      title = "Age"),
    
      barmode = "group"
    )
  

p1

```